---
title: "R and Rstudio"
author: "Fred Boehm"
date: "July 16, 2020"
output: html_document
---


## R 

- Command line-based environment for statistics and data analysis  
- Rich package collection  
  - Software  
  - Data  
- Diverse teaching materials in R  



## Rstudio

- User interface to R  
- Encourages script writing and other reproducible research practices  
- Enables git repository use  


## Rmarkdown

- Enhances (computational) reproducibility  
- One document holds both R code and narrative text
- Plain text format works well with `git`





## Example data analysis

```{r install, echo = TRUE}
remotes::install_github("allisonhorst/palmerpenguins")
```

```{r, echo = TRUE}
library(palmerpenguins)
library(ggplot2)
library(magrittr)
```


::: notes

These data were collected from 2007 - 2009 by Dr. Kristen Gorman with the [Palmer Station Long Term Ecological Research Program](https://pal.lternet.edu/), part of the [US Long Term Ecological Research Network](https://lternet.edu/). The data were imported directly from the [Environmental Data Initiative](https://environmentaldatainitiative.org/) (EDI) Data Portal, and are available for use by CC0 license ("No Rights Reserved") in accordance with the [Palmer Station Data Policy](https://pal.lternet.edu/data/policies). 

:::

## Data description

The curated `palmerpenguins::penguins` dataset contains `r ncol(penguins)` variables (n = `r nrow(penguins)` penguins):



::: notes

-   `species`: a factor denoting the penguin species (Adelie, Chinstrap, or Gentoo)
-   `island`: a factor denoting the island (in Palmer Archipelago, Antarctica) where observed (Biscoe, Dream, or Torgersen)
-   `bill_length_mm`: a number denoting length of the dorsal ridge of penguin bill (millimeters)
-   `bill_depth_mm`: a number denoting the depth of the penguin bill (millimeters)
-   `flipper_length_mm`: an integer denoting penguin flipper length (millimeters)
-   `body_mass_g`: an integer denoting penguin body mass (grams)
-   `sex`: a factor denoting penguin sex (male, female)

:::


```{r}
penguins[1:3, ]
```

## Exploratory data analysis

```{r mass-flipper}
gg <- ggplot(data = penguins, 
                       aes(x = flipper_length_mm,
                           y = body_mass_g)) +
  geom_point(aes(color = species, 
                 shape = species),
             alpha = 0.8)
gg
```

```{r}
plotly::ggplotly(gg)
```


## Exploratory data analysis

- Stratify by species, color by sex

```{r}
gg2 <- ggplot(penguins, aes(x = flipper_length_mm,
                            y = body_mass_g)) +
  geom_point(aes(color = sex)) +
  theme_minimal() +
  scale_color_manual(values = c("darkorange","cyan4"), na.translate = FALSE) +
  labs(title = "Penguin flipper and body mass",
       subtitle = "Dimensions for male and female Adelie, Chinstrap and Gentoo Penguins at Palmer Station LTER",
       x = "Flipper length (mm)",
       y = "Body mass (g)",
       color = "Penguin sex") +
  facet_wrap(~species)
gg2
```


```{r}
plotly::ggplotly(gg2)
```
